### 2020-07-11 analysis/paper/for paper ii.rmd

---
output:
  word_document: default
  pdf_document: default
  html_document: default
---

Some normal text.

```{r message=FALSE, warning=FALSE, echo=FALSE}
#install required packages
library(dplyr)
library(knitr)
library(ggplot2)
library(reshape2)
library(readxl)
library(fitdistrplus)
library(plyr)
library(tidyr)
library(classInt)
library(knitr)
library(cowplot)

# read in the data

file_name <- here::here("analysis/data/raw_data/artefacts-after Sam_Dec2018_quina_modified.xls")

flakes <- as.data.frame(read_excel(file_name, sheet = "flake basics"))
cores <- as.data.frame(read_excel(file_name, sheet = "core basics"))
debris <- as.data.frame(read_excel(file_name, sheet = "chunk&debris"))
retouch <- as.data.frame(read_excel(file_name, sheet = "retouch"))
core_flake_scars_1 <- as.data.frame(read_excel(file_name, sheet = "core flake scars"))
edge_angle_1 <- as.data.frame(read_excel(file_name, sheet = "edge angle"))
retouched_pebble_chunk <- as.data.frame(read_excel(file_name, sheet = "Pebble&chunk"))

scar_dir <- as.data.frame(read_excel(file_name, sheet = 'scar direction'))

core_flake_scars <- core_flake_scars_1[core_flake_scars_1$number %in% cores$number, ]

edge_angle <- edge_angle_1[edge_angle_1$number %in% retouch$number, ]

```

```{r message=FALSE, warning=FALSE, echo = FALSE}
# basic details of the dataset
artefact_ids <- list(flake_ids = flakes$number,
                     core_ids = cores$number,
                     debris_ids = debris$number)

retouch_ids <- retouch$number

# how many unique artefacts?

totals_of_each_type <- lapply(artefact_ids, function(i) length(i))

count_unique_artefacts <- sum(unlist(totals_of_each_type))

# debris pieces that are retouch pieces

debris_with_retouch <- intersect(retouch_ids, artefact_ids$debris_ids)

proportion_debris_retouched <- length(debris_with_retouch) / count_unique_artefacts

# proportion of each type

type_table <- data.frame(type = names(as.data.frame(totals_of_each_type)),
                         count = as.numeric(as.data.frame(totals_of_each_type)),
                         proportion = round(as.numeric(prop.table(as.data.frame(totals_of_each_type))),3))

# add total row at the bottom

type_table$type <- as.character(type_table$type)

type_table <- rbind(type_table, c("total", count_unique_artefacts, round(sum(type_table$proportion),2)))

# print pretty table

print("table of lithic types")

kable(type_table)

### raw materials

raw_material_table <- data.frame("category" = character(),
                                 'chert' = numeric(),
                                 'limestone' = numeric(),
                                 'basalt' = numeric(),
                                 'quartz' = numeric(),
                                 'sandstone' = numeric(),
                                 stringsAsFactors = FALSE)

row <- 1

raw_material_table[row, names(unlist(table(na.omit(flakes$material))))] <- 
  unlist(table(na.omit(flakes$material)))

raw_material_table[row, 1] <- "flake"
row <- row + 1

raw_material_table[row, names(unlist(table(na.omit(cores$material))))] <- 
  unlist(table(na.omit(cores$material)))
raw_material_table[row, 1] <- "core"
row <- row + 1

raw_material_table[row, names(unlist(table(na.omit(debris$material))))] <- 
  unlist(table(na.omit(debris$material)))
raw_material_table[row, 1] <- "debris"
row <- row + 1

raw_material_table[row, names(unlist(table(flakes$material[flakes$number %in% retouch$number])))] <- 
  unlist(table(flakes$material[flakes$number %in% retouch$number]))
raw_material_table[row, 1] <- "retouched pieces"
row <- row + 1

###tools raw materials

test <- as.matrix(t(table(flakes[flakes$number %in% retouch$number, 
                     c("material", "type2")])))

raw_material_table[row, colnames(test)] <- test["backed knife",]
raw_material_table[row, 1] <- "backed knife"
row <- row + 1

raw_material_table[row, colnames(test)] <- colSums( test[c("bec", "bec,scp"),])
raw_material_table[row, 1] <- "bec"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("borer","borer,dent", "borer,dent,scp","borer,end,scp","borer,notch",
                  "borer,notch,scp","borer,scp", "borer,scp,end","borer,strt" ), ])
raw_material_table[row, 1] <- "borer"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("BUR", "BUR? CORE?","BURIN" ), ])
raw_material_table[row, 1] <- "BUR"
row <- row + 1

raw_material_table[row, colnames(test)] <- test["chopper", ]
raw_material_table[row, 1] <- "chopper"
row <- row + 1

raw_material_table[row, colnames(test)] <- test["cleaver", ]
raw_material_table[row, 1] <- "cleaver"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("dent", "dent,borer,scp","dent,end","dent,notch",  "dent,scp",
                  "dent/NOT", "dent/POINT"  ), ])
raw_material_table[row, 1] <- "Denticular"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("end", "end,borer" , "end,borer,scp" ,"end,scp" ,"endscp" ), ])
raw_material_table[row, 1] <- "Endscraper"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("leva point","point" ,"POINT" , "point break" ,
                  "point,broken", "point?"  ), ])
raw_material_table[row, 1] <- "Point"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("moon","scp", "sidescp" , "sidescp-cvx"  ), ])
raw_material_table[row, 1] <- "scraper"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("NATBACK", "nature backed"), ])
raw_material_table[row, 1] <- "nature backed"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c( "notcch", "notch" , "notch,borer"  , "notch,end","notch,scp" ,       
                   "notch/DENT"), ])
raw_material_table[row, 1] <- "notch"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("TANG", "TANGED",  "TANGED POINT", "TANGED POINT /NOT", 
                  "tanged point?", "TANGED POINT?"  ), ])
raw_material_table[row, 1] <- "tanged point"
row <- row + 1

raw_material_table[is.na(raw_material_table)] <- 0 ##convert na to 0

raw_material_table$total <- rowSums(raw_material_table[,-1])

write.csv(raw_material_table, "raw_material_table.csv")

x <- raw_material_table

prop_raw_material <- cbind(category = x[, 1], round(x[, -1]/rowSums(x[, 2:(ncol(x)-1)])*100, digits = 1))

write.csv(prop_raw_material, "raw_material_table (percentage).csv")

print("tables for raw materials of the assemblage")

kable(raw_material_table)

print("proportion tables for raw materials of the assemblage")

kable(prop_raw_material)

### cores
core_summary <- cores[0, c(6:17)]

core_summary["mean",] <- c("mean" = apply(na.omit(cores[,c(6:17)]), 2, mean))

core_summary["sd",] <- c("sd" = apply(na.omit(cores[,c(6:17)]), 2, sd))

core_summary["CV",] <- core_summary["sd",]/core_summary["mean",]

core_summary <- rbind(core_summary, 
                      sapply(na.omit(cores[,6:17]), function(x) quantile(x, c(0.25, 0.5, 0.75))))

print("summary of core attributes")

kable(round(core_summary, digits = 1))

```

We will try univariate clustering for flake mass:

```{r message=FALSE, warning=FALSE, echo = FALSE}
require(Ckmeans.1d.dp)
result <- Ckmeans.1d.dp(flakes$mass[!is.na(flakes$mass)])

print("flake grouping result according to mass")
plot(result)

n_clusters <- length(result$size)
n_clusters_30 <- length(result$size[result$size > 30])

# assign cluster ID to table
flakes_clustered <- 
flakes %>% 
  filter(!is.na(mass)) %>% 
  mutate(cluster = result$cluster) %>% 
  right_join(flakes) %>% 
  filter(cluster %in% 1:n_clusters_30)

```

```{r message=FALSE, warning=FALSE, echo=FALSE}
# clean data to help with analysis
flakes$retouched <- ifelse(grepl("ret", flakes$type1), "retouched", "unretouched")
flakes <- flakes[!flakes$material %in% c('sandstone', 'quartz'),]

# how many levallois?
unique(flakes$type1)

leva <- flakes %>% 
  filter(grepl("LVT|LVF|LVFB", type1)) %>%
  filter(!grepl("LVF_", type1))

non_leva <- flakes %>% 
  filter(!grepl("LVT|LVF|LVFB", type1))

flakes_L <-  
  mutate(flakes, L = ifelse(grepl("LVT|LVF|LVFB", type1), "L", "N"))

# are the levallois pieces bigger?

print("compare the mass of Leva and non-leva flakes")
ggplot(flakes_L, aes(x = L, 
                     y = mass)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.2) +
  scale_y_log10() + theme_bw() +theme(text = element_text(size = 20))

# is it significant?

print("t-test to see if it is statistically significant?")

t.test(data = flakes_L, mass ~ L)
# no

# explore variation in thickness in these groups
thick_plot <- ggplot(flakes_L) +
  xlab("Thickness (mm)") + ylab("Density") +
  geom_density(aes(`Thickness_25%max`), colour = "red", alpha = 0.3) +
  geom_density(aes(`Thickness_50%max`), colour = "green", alpha = 0.3) +
  geom_density(aes(`Thickness_75%max`), colour = "blue", alpha = 0.3) +
  xlab("Thickness (mm)") + theme_bw() +
   theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

print("variation of thickness for all flakes (red = 25%max_dim,green = 50%max,blue=75%max")

thick_plot

# explore variation in width in these groups
width_plot <- ggplot(flakes_L) +
  xlab("Width (mm)") + ylab("Density") +
  geom_density(aes(`Width_25%max`), colour = "red", alpha = 0.3) +
  geom_density(aes(`Width_50%max`), colour = "green", alpha = 0.3) +
  geom_density(aes(`Width_75%max`), colour = "blue", alpha = 0.3) +
  xlab("Width (mm)") + theme_bw() +
   theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

print("variation of width for all flakes (red = 25%max_dim,green = 50%max,blue=75%max")

width_plot

# by size class that we identified earlier

print("compare thickness of leva and non-leva (red = 25%max_dim, 
      green = 50%max, blue=75%max")

thick_plot +
  facet_wrap(~ L, ncol = 1)

# need a test of uniformity (peak width)

# check the CVs
CV <- function(the_vector){
  mean_ <- mean(the_vector, na.rm = TRUE)
  sd_ <- sd(the_vector, na.rm = TRUE)
  cv <- (sd_/mean_) * 100
  round(cv, 3)
}

cv_res <- flakes_L %>% 
  group_by(L) %>% 
  dplyr::summarise(cv_25_thick = CV(`Thickness_25%max`),
                   cv_50_thick = CV(`Thickness_50%max`),
                   cv_75_thick = CV(`Thickness_75%max`),
                   cv_25_width = CV(`Width_25%max`), 
                   cv_50_width = CV(`Width_50%max`),
                   cv_75_width = CV(`Width_75%max`), 
                   n = n())

print("summary of cv for thickness and width for leva and non-leva flakes")

cv_res

```

The following are results for cores

```{r message=FALSE, warning=FALSE, echo=FALSE}

# function to compute frequencies and make a plot
f <-  function(the_data, the_column) {
  data.frame(table(the_data[the_column])) %>% 
    ggplot(aes(x = reorder(Var1, Freq), y = Freq)) +
    geom_bar(stat = "identity") +
    ylab("count") +
    xlab(the_column) + 
    theme_bw() +
    theme(axis.text.x = element_text(size = 20, angle = 45, hjust = 1, vjust = 1),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25)
    )
}

# some plots of frequencies for cores

## geometry
print("summary for core geometry")

f(cores, "core_geometry")

core_geom_table <-  table(na.omit(cores$core_geometry))

round(prop.table(core_geom_table), digits = 4)*100

## platform number

print("summary for core platform")

f(cores, "platform")

core_plat_table <-  table(na.omit(cores$platform))

round(prop.table(core_plat_table), digits = 4)*100

### cortex proportion of cores

print("cortex proportion of cores")

ggplot(cores, aes(cortex_percentage)) +
  geom_histogram(binwidth = 10) + theme_bw() +
  xlab("Cortex proportion (%)") + 
  ylab("Number of cores") +
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

core_cortex_table <-  table(na.omit(cores$cortex_percentage))

round(prop.table(core_cortex_table), digits = 4)*100

## platform type

print("summary for core platform preparation")

f(cores, "platform_preparation")

core_plat_type <-  table(na.omit(cores$platform_preparation))

round(prop.table(core_plat_type), digits = 4)*100

## platform rotation
print("summary for core rotations")

f(cores, "rotations")

ggplot(cores, aes(rotations)) +
  geom_bar() + theme_bw() +
  xlab("Number of rotation") + 
  ylab("Number of cores") +
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

core_plat_rotat <-  table(na.omit(cores$rotations))

round(prop.table(core_plat_rotat), digits = 4)*100

## core reduction type
print("summary for core reduction type")

f(cores, "type")

core_type_table <- table(na.omit(cores$type))

round(prop.table(core_type_table),digits = 4)*100

## core scar number

print("summary of core scar number")

ggplot(cores, aes(scar_number)) +
  geom_bar() + theme_bw() + 
  xlab("Scar number on cores") + 
  ylab("Count") +
    theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

core_scar_table <- table(na.omit(cores$scar_number))

round(prop.table(core_scar_table), digits = 4)*100

## core flake scar length

print("summary for core flake scar length")

core_flake_scars_melt <- melt(core_flake_scars_1)

colnames(core_flake_scars_melt)[3] <- "scar_length"

ggplot(core_flake_scars_melt[core_flake_scars_melt$scar_length < 500, ], 
       aes(scar_length)) +
  geom_density() + theme_bw() +
  xlab("Length of scar on cores") + 
  ylab("Density") +
    theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

print("median of scar length")

median(na.omit(core_flake_scars_melt[core_flake_scars_melt$scar_length < 500, ]$scar_length))

```

The following results are for flakes

```{r message=FALSE, warning=FALSE, echo=FALSE}

# explore flakes

flake_summary <- flakes[0, c(7:17, 19,20,22,23)]

flake_summary["mean",] <- c("mean" = apply(na.omit(flakes[,c(7:17, 19,20,22,23)]), 2, mean))

flake_summary["sd",] <- c("sd" = apply(na.omit(flakes[,c(7:17, 19,20,22,23)]), 2, sd))

flake_summary["cv", ] <- flake_summary["sd",] /flake_summary["mean",]

flake_summary <- rbind(flake_summary, 
                      sapply(na.omit(flakes[,c(7:17, 19,20,22,23)]), 
                             function(x) quantile(x, c(0.25, 0.5, 0.75))))

print("table for flake attributes")

kable(round(flake_summary, digits = 1))

## how many flake type1

print("summary for flake type1")
f(flakes, "type1")
table(na.omit(flakes$type1))
round(prop.table(table(na.omit(flakes$type1))), digits = 3)*100

## flake type2 
print("summary for flake type2")
f(flakes, "type2")
table(na.omit(flakes$type2))
round(prop.table(table(na.omit(flakes$type2))),digits = 3)*100

## how many retouched and unretouched flakes
flakes$retouched <- ifelse(grepl("ret", flakes$type1), "retouched", "unretouched")

flakes <- flakes[!flakes$material %in% c('qutz', 'quartz', 'basalt'),]

print("summary for flake retouch")

f(flakes, "retouched")
table(na.omit(flakes$retouched))
round(prop.table(table(na.omit(flakes$retouched))), digits = 3)*100

## raw material of flakes
print("summary for flake material")
f(flakes, "material")

## platform shape of flakes
print("summary for flake platform shape")
f(flakes, "platform_shape")

table(na.omit(flakes$platform_shape))
round(prop.table(table(na.omit(flakes$platform_shape))), digits = 4)*100

### number of flakes with distinctive platform
## platform of flakes
print("summary for flake platform")
f(flakes, "platform")

paste( "number of flakes with distinctive platform =", length(na.omit(flakes$platform)) )

table(na.omit(flakes$platform))
round(prop.table(table(na.omit(flakes$platform))), digits = 3)*100

# mass for all flakes
print("mass of flakes")
ggplot(flakes, aes(mass)) +
  geom_histogram(binwidth = 10) + theme_bw() +
  xlab("Mass (g)") + ylab("Number of flakes") +
    theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

## max dimension of all flakes

print("max dimension of all flakes")
ggplot(flakes, aes(max_dimension)) +
  geom_density() + theme_bw() +
  xlab("Maximum dimension (mm)") + ylab("Density") +
    theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

## length vs width at 50% max dime
print("length vs width at 50% max dim")

ggplot(flakes, aes(length, `Width_50%max`, colour = material)) +
  geom_point() +
  geom_smooth(method = "lm") + theme_bw() +
    theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

## platform shape vs thickness at 50% max dime
print("platform shape vs thickness at 50% max dim")

ggplot(flakes[!is.na(flakes$platform_shape),], 
       aes(`Thickness_50%max`, x = platform_shape)) +
  geom_boxplot() + scale_y_log10()+ theme_bw() +
  xlab("Platform shape") + ylab("Thickness at 50% max. dim. (mm)") +
    theme(axis.text.x = element_text(size = 20, angle = 45, vjust = 1, hjust = 1),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

## platform shape vs max dim
print("platform shape vs max dim")

ggplot(flakes[!is.na(flakes$platform_shape),],
       aes(`max_dimension`, x = platform_shape)) +
  geom_boxplot() + scale_y_log10()+ theme_bw() +
   xlab("Platform shape") + ylab("Maximum dimension (mm)") +
    theme(axis.text.x = element_text(size = 20, angle = 45, vjust = 1, hjust = 1),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

# mass for all flakes, retouched or not

print("mass for retouched and non-retouched flakes")

ggplot(flakes, aes(mass, colour = retouched)) +
  geom_density()+ theme_bw() +

# mass for all flakes, by raw material

print("mass for flakes of different raw materials")
ggplot(flakes, aes(mass, colour = material)) + scale_x_log10() +
  geom_density()+ theme_bw() +theme(text = element_text(size = 20))

## max dimension for all flakes

print("max dim for all flakes")
ggplot(flakes, aes(max_dimension)) +
  geom_histogram(binwidth = 5) + theme_bw() +
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

## max dim for all flakes, retouched or not

print("max dim for all flakes, retouched or not")
ggplot(flakes, aes(`max_dimension`, colour = retouched)) +
  geom_density()+ theme_bw() +
  xlab("Maximum dimension (mm)") + ylab("Density") +
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25),
        legend.text = element_text(size = 20))+
   theme(legend.position = c(0.8, 0.8),
         legend.title = element_blank())

## max dim for all flakes, by raw material
print("max dim for all flakes, by raw material")
ggplot(flakes, aes(`max_dimension`, colour = material)) +
  geom_density()+ theme_bw() +theme(text = element_text(size = 20))

## scar counts for all flakes, by raw material
print("scar counts for all flakes, by raw material")
ggplot(flakes, aes(`scar_number`)) +
  geom_bar()+ theme_bw() +
  xlab("Number of scars") + ylab("Count of flakes")+
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

flakes$scar_group <- ifelse(flakes$`scar_number` > 5, "more_than_5", "less_than_5")

ggplot(flakes, aes(mass, x = scar_group)) +
  geom_boxplot() +
  scale_y_log10()+ theme_bw() +theme(text = element_text(size = 20))

ggplot(flakes[!is.na(flakes$scar_group), ], aes(mass, colour = scar_group)) +
  geom_density()+ theme_bw() + scale_x_log10() +
  xlab("Mass (g)") + ylab("Count of flakes")+
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

# facetted plaforms

print("facetted platforms")

flakes$facetted_platform <- ifelse(grepl("fac", flakes$platform), "fac", "no")

table(na.omit(flakes$facetted_platform))
round(prop.table(table(na.omit(flakes$facetted_platform))), digits = 4)*100

ggplot(flakes, aes(`max_dimension`, colour = facetted_platform)) +
  geom_density()+ theme_bw() +
  xlab("Maximum dimension (mm)") + ylab("Density") +
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25),
        legend.text = element_text(size = 20))+
   theme(legend.position = c(0.8, 0.8),
         legend.title = element_blank())

### cortex proportion of flakes

print("cortex proportion of flakes")

ggplot(flakes, aes(cortex_percentage)) +
  geom_histogram(binwidth = 10) + theme_bw() +
    xlab("Cortex proportion (%)") + ylab("Count of flakes") +
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

### compare cortex for ret and non-ret

print("compare cortex for ret and non-retouched flakes")

ggplot(flakes, aes(x = cortex_percentage, fill = retouched)) + geom_histogram(binwidth = 10, position='dodge') + theme_bw() +theme(text = element_text(size = 20))

## how many points

print("how many points?")

flakes$point <- ifelse(grepl("point|POINT", flakes$type2), "point", 
                       ifelse(grepl("back", flakes$type2), "backed", "other"))

table(na.omit(flakes$point))
round(prop.table(table(na.omit(flakes$point))), digits = 4)*100

```

Group the flakes according to their

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Look to see if groups exist

# by flake size
plain_flakes <-  flakes[grepl("fl", flakes$type1), ]
d <- density(na.omit(plain_flakes$mass)) # returns the density data
# plot
plot(d)
ggplot(plain_flakes, aes(mass)) +
  geom_density()+ theme_bw() +theme(text = element_text(size = 20))

# test is we can discover natural breaks that divide up the flakes?

# how many groups should we look for?
n <- 5
natural_breaks <- classIntervals(plain_flakes$mass, n = n, 
                                 style = "jenks", dataPrecision = 2)
# plot the breaks
ggplot(plain_flakes, aes(mass)) +
  geom_density(colour = "blue") +
  geom_vline(xintercept = natural_breaks$brks, colour = "red")+ theme_bw() +theme(text = element_text(size = 20))

# create a df
plain_flakes$mass_group <- cut(plain_flakes$mass, 
                               breaks = c(0, natural_breaks$brks[-1]),
                               labels = 1:(length(natural_breaks$brks)-1))
# remove flakes not in a group
plain_flakes <- plain_flakes[!is.na(plain_flakes$mass_group), ]

# plot the groups
ggplot(plain_flakes, aes(mass_group, mass)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.1) +
  theme_bw() +theme(text = element_text(size = 20))

```

```{r message=FALSE, warning=FALSE, echo=FALSE}
# explore variation in thickness in these groups
thick_plot <- ggplot(plain_flakes) +
  xlab("Thickness (mm)") + ylab("Density") +
  geom_density(aes(`Thickness_25%max`), colour = "red", alpha = 0.3) +
  geom_density(aes(`Thickness_50%max`), colour = "green", alpha = 0.3) +
  geom_density(aes(`Thickness_75%max`), colour = "blue", alpha = 0.3) +
  xlab("Thickness (mm)") + theme_bw() +theme(text = element_text(size = 20))

thick_plot

# by size class that we identified earlier
thick_plot +
  facet_wrap(~ mass_group, ncol = 1)

# Lycett and Eren found out the thickness is more evenly distributed and less variable across preferential Levallois flakes. So , let's get the CV values for thickness for each group
CV <- function(the_vector){
  mean_ <- mean(the_vector, na.rm = TRUE)
  sd_ <- sd(the_vector, na.rm = TRUE)
  cv <- (sd_/mean_)*100
  round(cv, 3)
}

plain_flakes %>% 
  group_by(mass_group) %>% 
  dplyr::summarise(cv_25_thick = CV(`Thickness_25%max`),
                   cv_50_thick = CV(`Thickness_50%max`),
                   cv_75_thick = CV(`Thickness_75%max`),
                   cv_25_width = CV(`Width_25%max`), 
                   cv_50_width = CV(`Width_50%max`),
                   cv_75_width = CV(`Width_75%max`), 
                   n = n())

```

Results for edge angles

```{r message=FALSE, warning=FALSE, echo=FALSE}
# calculate the edge angle

edge_angle_2 <- edge_angle

edge_angle_2[, 2:(ncol(edge_angle_2)-1)] <- atan(edge_angle[, 2:(ncol(edge_angle_2)-1)]/2/3)/pi*180*2

df_angle <- melt(edge_angle_2[, 1:(ncol(edge_angle_2)-1)])

df_angle$variable <- substr(df_angle$variable, 1, 9)

colnames(df_angle)[2:3] <- c("section", "angle")

# plot the angles

print("edge angels")

ggplot(df_angle, aes(x = angle, col = section)) + geom_density() + theme_bw() +theme(text = element_text(size = 20))

## find quina and compare their angles.

quina_id <- flakes$number[which(flakes$type3 == "quina")]

df_angle$quina <- ifelse(df_angle$number %in% quina_id, "quina", "no")

ggplot(df_angle, aes(x = angle, col = quina)) + geom_density() + theme_bw() +theme(text = element_text(size = 20))

ggplot(df_angle, aes(x = angle, col = section)) + geom_density() + theme_bw() +theme(text = element_text(size = 20)) + facet_wrap(~quina)

## compare thickness of quina and other

df_thick <- melt(flakes[, c("number", "Thickness_25%max", "Thickness_50%max", "Thickness_75%max")])

df_thick$quina <- ifelse(df_thick$number %in% quina_id, "quina", "no")

ggplot(df_thick, aes(x = value, col = quina)) + geom_density() + theme_bw() +theme(text = element_text(size = 20)) + facet_wrap(~variable)

# plot the thickness of 50% max

print("thickness of 50% max dim")
ggplot(df_thick[df_thick$variable =="Thickness_50%max", ], aes(quina, value)) +
  geom_boxplot() + theme_bw() +theme(text = element_text(size = 20))

## count the number of shapes of edges.

edge_shapes <- unlist(strsplit(retouch$`edge shape`, ","))

data.frame(table(edge_shapes)) %>% 
ggplot(aes(x = reorder(edge_shapes, Freq), y = Freq)) +
    geom_bar(stat = "identity") +
    ylab("n") +
    xlab("Edge type") + 
    theme_bw() +theme(text = element_text(size = 20)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

data.frame(table(edge_shapes))

## number of layers

f(retouch, 'number of layers')

## number of edges

f(retouch, 'number of edge')

## number of layers for each number of eadge

df_layer_edge <- melt(retouch[, 1:3])

 data.frame(table(retouch[,2:3])) %>% 
    ggplot(aes(x = reorder(number.of.layers, Freq), y = Freq)) +
    geom_bar(stat = "identity") +
    ylab("n") +
    xlab("number of layers") + 
    theme_bw() +theme(text = element_text(size = 20)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    facet_wrap(~number.of.edge)

```

scar number and direction

```{r message=FALSE, warning=FALSE, echo=FALSE}

# scar number and direction

## count how many scars from each direction

print("count how many scars from each direction")

table(unlist(scar_dir[,2:8]))

## How many % flakes that have dorsal scars from at least 3 complete different directions

print("How many % flakes that have dorsal scars from at least 3 complete different directions")

table(unlist(scar_dir[,2:8]))

## calculate GIUR

df_giur <- retouch[, c(1,8,10,12,14,16,18,20,22)]

df_giur[, -1] <- df_giur[, -1] / retouch[, c(9,11,13,15,17,19,21,23)]

df_giur_clustered <- left_join(df_giur, flakes_clustered[, c("number", "cluster")])

df_giur_clustered_melt <- melt(df_giur_clustered[,-1], id.vars = "cluster")

## compare GIUR with edge angle

df_giur_edge <- melt(df_giur)

df_giur_edge$variable <- substr(df_giur_edge$variable, 1, 9)

colnames(df_giur_edge)[2:3] <- c("section", "giur")

df_giur_edge <- left_join(df_giur_edge[df_giur_edge$giur <= 1, ], df_angle)

print("compare the GIUR and edge angle for different sections")

ggplot(na.omit(df_giur_edge), aes(angle, giur, colour = section)) +
  geom_point() +
  geom_smooth(method = "lm") + theme_bw() +theme(text = element_text(size = 20))

print("compare the GIUR and edge angle for all sections")

ggplot(na.omit(df_giur_edge), aes(angle, giur)) +
  geom_point() +
  geom_smooth(method = "lm") + theme_bw() +theme(text = element_text(size = 20))

## compare GIUR for different mass categories

print("compare GIUR for different mass groups")

ggplot(na.omit(df_giur_clustered_melt[df_giur_clustered_melt$value <=1,]), aes(value)) +
  geom_histogram() + theme_bw() +theme(text = element_text(size = 20)) + facet_wrap(~cluster)

## calculate the median GIUR for each group

print("the median GIUR for each group")

tapply(df_giur_clustered_melt$value, df_giur_clustered_melt$cluster, median, na.rm=TRUE)

## compare GIUR for Quina and non-quina

print("Compare GIUR for Quina and non-quina")

df_giur_clustered$quina <- ifelse(df_giur_clustered$number %in% quina_id, "quina", "no")

df_giur_clustered_melt_2 <- melt(df_giur_clustered[,-1], id.vars = "quina")

ggplot(na.omit(df_giur_clustered_melt_2[df_giur_clustered_melt_2$value <=1,]), aes(value)) +
  geom_histogram() + theme_bw() +theme(text = element_text(size = 20)) + facet_wrap(~quina)

print("Histogram of GIUR for Quina")

ggplot(na.omit(df_giur_clustered_melt_2[df_giur_clustered_melt_2$value <=1 &
                                          df_giur_clustered_melt_2$quina == "quina",]), aes(value)) +
  geom_histogram()  + theme_bw() +theme(text = element_text(size = 20))

print("Histogram of GIUR for non Quina")

ggplot(na.omit(df_giur_clustered_melt_2[df_giur_clustered_melt_2$value <=1 &
                                          df_giur_clustered_melt_2$quina == "no",]), aes(value)) +
  geom_histogram() +  theme_bw() +theme(text = element_text(size = 20))

print("the median GIUR for quina and non-quina")

tapply(df_giur_clustered_melt_2$value, df_giur_clustered_melt_2$quina, median, na.rm=TRUE)

# Invasiveness index
print("results for invasiveness index")

df_invasive <- retouch[, c(1, 24:39)]

df_invasive_clustered <- left_join(df_invasive, flakes_clustered[, c("number", "cluster")])

df_invasive_clustered_melt <- melt(df_invasive_clustered[,-1], id.vars = "cluster")

invasive_table <- as.data.frame(table(df_invasive_clustered_melt[, c("cluster", "value")]))

ggplot(invasive_table[invasive_table$value %in% c(0.5, 1.0), ], aes(fill=value, y= Freq, x=cluster)) + 
    geom_bar( stat="identity", position="fill", width = 0.6) + theme_bw() +theme(text = element_text(size = 20)) +
    theme(axis.text=element_text(size=14), axis.title=element_text(size=16,face="bold")) +
    theme(legend.text=element_text(size=14))

```