### 2020-10-04 analysis/paper/old-stuff/flake_attributes_and_size.Rmd
tidy up and polish fig 1

---
title: "Guanyingdong Stone Artefact Assemblage Report: Variation by Size"
author: "HY and BM"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   bookdown::word_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
---

```{r setup-flake-size, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Nature’s standard figure sizes are 89 mm (single column) and 183 mm (double column)
 # https://www.nature.com/nature/for-authors/formatting-guide
```

THe reviewers ask:

>  can they offer other data that might elucidate changes in flaking type with reduction, e.g. changes in flake attributes that correlate with size declines?

Read in the data, deal with dupicates, get the basic flake details, run to L454 of hu_artefacts_report.Rmd, then continue from here.

Let's see if there are any natural groupings in flake mass:

```{r}

flake_hist_1 <- 
ggplot(flakes,
       aes(mass)) +
  geom_histogram() +
  theme_minimal() +
  xlab("Mass (g)")

flake_hist_2 <- 
ggplot(flakes,
       aes(mass)) +
  geom_histogram() +
  theme_minimal() +
  scale_x_log10() +
  xlab("Mass (g)")

library(cowplot)

flake_mass_distributions <- plot_grid(flake_hist_1, flake_hist_2, ncol = 1)

```

Nothing obvious, let's try a little clustering. We will try univariate clustering:

```{r}
require(Ckmeans.1d.dp)
result <- Ckmeans.1d.dp(flakes$mass[!is.na(flakes$mass)])
plot(result)

n_clusters <- length(result$size)
n_clusters_30 <- length(result$size[result$size > 30])

# assign cluster ID to table
flakes_clustered <- 
flakes %>% 
  filter(!is.na(mass)) %>% 
  mutate(cluster = result$cluster) %>% 
  right_join(flakes) %>% 
  filter(cluster %in% 1:n_clusters_30)
```

We get `r n_clusters` clusters, with `r n_clusters_30` that have more than 30 artefacts in them

```{r}
ggplot(flakes_clustered,
       aes(mass, 
          length,
          colour = factor(cluster)))+
  geom_point()

flake_hist_with_clusters <- 
ggplot(flakes_clustered,
       aes(mass)) +
  geom_histogram(aes(fill = factor(cluster))) +
  annotate("text", x = 240, y = 120, label = glue::glue('Total number of artefacts in sample: {nrow(flakes_clustered)}')) +
  scale_fill_viridis_d(name = "Size class") +
  theme_minimal() +
  xlab("Mass (g)")  +
  coord_equal(ratio = 0.2) +
  theme(legend.position = c(.75, .75),
        legend.direction = "horizontal",
        legend.key.size = unit(0.6,"line"),
        legend.key.width = unit(0.6,"line"),
        legend.key.height = unit(0.6,"line"),
        legend.text = element_text(size=8),
        legend.box.background = element_rect(fill = "white", colour = "white"))
```

Now let's see how some of the key variables change accross these size clusters:

We see that some dimensions don't increase with size. Platform thickness decreases a little, as mass goes up, platform width becomes more variable. Scar numbers have less high outliers as size increases.

```{r}

library(stringr)
# get sample sizes
flakes_clustered_numeric_vars_sample_sizes <- 
flakes_clustered %>% 
  select_if(is.numeric) %>%
  gather(variable, value, -cluster) %>% 
  mutate(variable = str_to_title(str_replace_all(variable, "_", " "))) %>% 
  filter(!is.na(value)) %>% 
  mutate(variable = str_replace_all(variable, "%", "% ")) %>% 
  mutate(cluster = factor(cluster)) %>% 
  group_by(variable) %>% 
  tally()

# add measurement units
flakes_clustered_numeric_vars_sample_sizes$units <- 
  c("", 
    "mm, ",
    "g, ",
    "mm, ",
    "mm, ",
    "mm, ",
    "mm, ",
    "mm, ",
    "",
    "mm, ",
    "mm, ",
    "mm, ",
    "mm, ",
    "mm, ",
    "mm, ")

flakes_clustered_numeric_vars_plot <- 
flakes_clustered %>% 
  select_if(is.numeric) %>%
  gather(variable, value, -cluster) %>% 
  mutate(variable = str_to_title(str_replace_all(variable, "_", " "))) %>% 
  filter(!is.na(value)) %>% 
  mutate(variable = str_replace_all(variable, "%", "% ")) %>% 
  mutate(cluster = factor(cluster)) %>% 
  left_join(flakes_clustered_numeric_vars_sample_sizes) %>% 
  # mutate(variable = glue::glue('{variable} ({units}n = {n})')) %>% 
  ggplot(aes(cluster,
             value)) +
  geom_boxplot() +
  facet_wrap( ~ variable, scales = "free_y") +
  xlab("Size class") +
  ylab("") +
  theme_minimal(base_size = 8)

figure_one <- 
cowplot::plot_grid(flake_hist_with_clusters, 
                   flakes_clustered_numeric_vars_plot,
                   labels = "AUTO", 
                   ncol = 1, 
                   align = 'v', 
                   axis = 'lr', 
                   rel_heights = c(0.5, 1.5)) #

save_plot("Extended_Data_Figure_X.png", figure_one,
          base_aspect_ratio = 1, 
          base_height = 8,
          base_width = 18.3/2.5)
```

Platform thickness decreases, width at 50% and 75% does not increase as much as width at 25% suggesting a slight increase in triangular-shaped flakes over parallel-sided flakes as the flakes get larger.

How about the categorical variables?

```{r}

flakes_categorical <- 
flakes_clustered %>% 
  select_if(function(col) is.character(col) | 
                                   all(col == .$cluster)) %>% 
  gather(variable, value, -cluster) %>% 
  mutate(cluster = factor(cluster)) %>% 
  filter(!is.na(value)) %>% 
  group_by(cluster, variable, value) %>% 
  tally() %>% 
  ungroup() %>% 
  group_by(cluster, variable) %>%  
  mutate(prop = n / sum(n) * 100)

# tidy up some of these variables
library(stringr)
library(forcats)

flakes_clustered_cats <- 
flakes_clustered %>% 
  select_if(function(col) is.character(col) | 
                                   all(col == .$cluster))

# raw material -----------------------------------------------------------------
# material, ok
flakes_clustered_cats %>% 
 group_by(material) %>% 
 tally(sort = TRUE)

# make table for plotting
flakes_clustered_cats_raw_material <- 
flakes_clustered_cats %>%  
 group_by(`material` , cluster) %>% 
 tally(sort = TRUE)

# flake type -----------------------------------------------------------------

# type1, ok
flakes_clustered_cats %>% 
 group_by(type1) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_flake_types <- 
  flakes_clustered_cats %>% 
  mutate(type1 = if_else(str_detect(type1, "lvf"), "Levallois", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "ret"), "retouch", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "blade|bld"), "blade", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "flk|flake"), "flake", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "dbd"), "débordant", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "kbw"), "Kombewa", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "end"), "end scraper", type1)) %>% 
  mutate(type1 = fct_lump(type1, 5))

# check again
flakes_clustered_cats_flake_types %>% 
 group_by(type1) %>% 
 tally(sort = TRUE)

# for plotting
flakes_clustered_cats_flake_types <- 
flakes_clustered_cats_flake_types %>% 
 group_by(`type1`, cluster) %>% 
 tally(sort = TRUE)

# retouch type -----------------------------------------------------------------

# type 2, special case , retouch types
flakes_clustered_cats %>% 
 group_by(type2) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_retouch_type <- 
flakes_clustered_cats %>% 
 separate(type2, into = c("type2a", "type2b", "type2c"), sep = ",") %>% 
  mutate(type2a = if_else(str_detect(type2a, "lvf"), "levallois", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "ret"), "retouch", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "blade|bld"), "blade", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "flk|flake"), "flake", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "tang"), "tanged", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "not"), "notch", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "bur"), "burin", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "point"), "point", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "dent"), "denticulate", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "scp"), "scraper", type2a)) %>% 
  mutate(type2a = str_squish(type2a))  %>% 
  mutate(type2a = fct_lump(type2a, 7)) %>% 
  filter(!is.na(type2a))

# check it
flakes_clustered_cats_retouch_type %>% 
 group_by(type2a) %>% 
 tally(sort = TRUE)

# for plotting
flakes_clustered_cats_retouch_type <- 
flakes_clustered_cats_retouch_type %>% 
# rename(`Retouch type` = type2a ) %>% 
# group_by(`Retouch type`, cluster) %>% 
 tally(sort = TRUE)

# platform type -----------------------------------------------------------------

# Platform, ok
flakes_clustered_cats %>% 
 group_by(platform) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_platform_type  <- 
  flakes_clustered_cats %>% 
   mutate(platform = if_else(str_detect(platform, "facet"), "facetted", platform)) %>% 
   mutate(platform = if_else(str_detect(platform, "cotex|cortex"), "cortex", platform)) %>%
   mutate(platform = fct_lump(platform, 6)) %>% 
  filter(!is.na(platform))

# for plotting
flakes_clustered_cats_platform_type <- 
flakes_clustered_cats_platform_type %>% 
 group_by(`platform`, cluster) %>% 
 tally(sort = TRUE)

# platform shape -----------------------------------------------------------------

# platform shape, ok
flakes_clustered_cats %>% 
 group_by(platform_shape) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_platform_shape <- 
flakes_clustered_cats %>% 
  mutate(platform_shape = if_else(platform_shape == "sb", "quadrangle", platform_shape)) %>%
  mutate(platform_shape = if_else(platform_shape == "cdg", "chapeau de gendarme", platform_shape)) %>%
  mutate(platform_shape = if_else(platform_shape == "sx", "rhombus", platform_shape)) %>%
  mutate(platform_shape = fct_lump(platform_shape, 4))  %>% 
  filter(!is.na(platform_shape))

# for plotting
flakes_clustered_cats_platform_shape <- 
flakes_clustered_cats_platform_shape %>% 
# rename(`platform` = platform_shape ) %>% 
 group_by(`platform_shape`, cluster) %>% 
 tally(sort = TRUE)

# cortex texture-----------------------------------------------------------------

# cortex texture, ok 
flakes_clustered_cats %>% 
 group_by(cortex_texture) %>% 
 tally(sort = TRUE)

# for plotting
flakes_clustered_cats_cortex_texture <- 
flakes_clustered_cats %>% 
 group_by(`cortex_texture`, cluster) %>% 
 filter(!is.na(`cortex_texture` )) %>% 
 tally(sort = TRUE)

# cortex location-----------------------------------------------------------------

# cortex location, ok
flakes_clustered_cats %>% 
 group_by(cortex_location) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_cortex_location <- 
  flakes_clustered_cats %>% 
   mutate(cortex_location = if_else(str_detect(cortex_location, "prox"), 
                                    "proximal", 
                                    cortex_location)) %>%
   mutate(cortex_location = fct_lump(cortex_location, 6)) %>% 
  filter(!is.na(cortex_location),
         cortex_location != "vent")

flakes_clustered_cats_cortex_location <- 
flakes_clustered_cats_cortex_location %>% 
 group_by(`cortex_location`, cluster) %>% 
 filter(!is.na(`cortex_location` )) %>% 
 tally(sort = TRUE)

# cortex location-----------------------------------------------------------------

# initiation, ok 
flakes_clustered_cats %>% 
 group_by(initiation) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_initiation <- 
  flakes_clustered_cats %>% 
   mutate(initiation = fct_lump(initiation, 3))

flakes_clustered_cats_initiation <- 
flakes_clustered_cats_initiation %>% 
 group_by(`initiation`, cluster) %>% 
 filter(!is.na(`initiation` )) %>% 
 tally(sort = TRUE)

# cortex location-----------------------------------------------------------------

# platform preparation, ok, ignore, no meaningful info in here
flakes_clustered_cats %>% 
 group_by(platform_preparation) %>% 
 tally(sort = TRUE)

flakes_clustered_cats <- 
  flakes_clustered_cats %>% 
   mutate(platform_preparation = if_else(str_detect(platform_preparation, "fac"), "facetted", platform_preparation)) %>% 
   mutate(platform_preparation = fct_lump(platform_preparation, 3))

```

Make a nice panel plot, we have separate dataframes for each variable.

```{r}
list_of_dfs <- list(
  #flakes_clustered_cats_retouch_type,
                    flakes_clustered_cats_raw_material,
                    flakes_clustered_cats_platform_type,
                    flakes_clustered_cats_platform_shape,
                    flakes_clustered_cats_initiation,
                    flakes_clustered_cats_flake_types,
                    flakes_clustered_cats_cortex_texture,
                    flakes_clustered_cats_cortex_location)

library(purrr)

map_dfr(list_of_dfs, ~nest(.x))

all_the_categorical_variables <- 
  map_dfr(list_of_dfs, 
          ~gather(.x, variable, value, -cluster, -n))

# Compute proportions per category

flakes_clustered_cats_long <- 
all_the_categorical_variables %>% 
  mutate(value = str_squish(value)) %>% 
  mutate(cluster = factor(cluster)) %>% 
  group_by(cluster, variable) %>%  
  mutate(prop = n / sum(n) * 100)  %>% 
  arrange(cluster, variable)
```

Here is the plot

```{r}
# sample sizes for each variable
flakes_clustered_cats_sample_sizes <- 
flakes_clustered_cats_long %>% 
  group_by(variable) %>% 
  summarise(tot_n = sum(n))

categorical_plots <- 
flakes_clustered_cats_long %>% 
  ungroup(cluster) %>% 
  nest(-variable) %>% 
  left_join(flakes_clustered_cats_sample_sizes) %>% 
  mutate(variable = glue::glue('{variable} (n = {tot_n})')) %>% 
  mutate(data = map(data, ~.x %>% mutate(value = as.character(value)))) %>% 
  mutate(plot = map2(data,
                     variable, ~.x %>%  
                            ggplot(aes(cluster, 
                                          prop, 
                                          fill = value)) + 
                       scale_fill_viridis_d() +
                            geom_col() +
                            xlab("") +
                            ylab("%") +
                       theme_minimal(base_size = 8) +
                           theme(legend.key.size = unit(0.4,"line"),
                                 legend.key.width = unit(0.4,"line"),
                                 legend.key.height = unit(0.4,"line"),
                                 legend.text = element_text(size=8), 
                                 legend.title = element_blank(),
                                 legend.margin=margin(t=-0.5, 
                                                      r=-0.5, 
                                                      b=-0.5, 
                                                      l=-0.1, 
                                                      unit="cm"),
                                 legend.box.margin=margin(-0.1,
                                                          -0.1,
                                                          -0.1,
                                                          -0.3,
                                                          unit="cm"),
                                 plot.margin = unit(x = c(0.5, 
                                                          0, 
                                                          0,
                                                          0.5), 
                                                    units = "cm")) +
                            ggtitle(str_glue('{.y}'))))

# arrange and draw the plots
categorical_plots_arranged <- 
cowplot::plot_grid(plotlist = categorical_plots$plot,
                   align = "hv",
                   axis = "tblr",
                   ncol = 2)

save_plot("analysis/report/Extended_Data_Figure_Y.png", categorical_plots_arranged,
          base_aspect_ratio = 1,
          base_height = 8,
          base_width = 18.3/2.5
)

```

A closer look at the distribution of retouch types by size classes

```{r eval=FALSE}
flakes_clustered_cats_long %>% 
  filter(variable == "Retouch type") %>% 
  ggplot(aes(value, prop)) +
  geom_col() +
  facet_wrap( ~ cluster , ncol = 1)
```

A closer look at the distribution of retouch types by size classes

```{r eval=FALSE}
flakes_clustered_cats_long %>% 
  filter(variable == "Retouch type") %>% 
  ggplot(aes(value, prop)) +
  geom_col() +
  facet_wrap( ~ cluster , ncol = 1)
```

Try this on the Lev flakes only:

```{r  eval=FALSE}
flakes_clustered %>% 
  select_if(function(col) is.numeric(col) | 
                                   all(col == .$type1))  %>% 
  filter(str_detect(type1, "LV|leva")) %>% 
    select(-box, -type1) %>% 
  gather(variable, value, -cluster) %>% 
  filter(!is.na(value)) %>% 
  mutate(cluster = factor(cluster)) %>% 
  ggplot(aes(cluster,
             value)) +
  geom_boxplot() +
  facet_wrap( ~ variable, scales = "free_y") +
  theme_minimal()
```

```{r  eval=FALSE}
flakes_clustered_cats_long_leva <- 
flakes_clustered_cats %>% 
  filter(type1 %in% c("levallois", "LVF")) %>% 
  gather(variable, value, -cluster)  %>% 
  filter(!is.na(value), 
         variable != "comment",
         variable != "heat_damage",
         variable != "platform_preparation",
         variable != "type2c") %>% 
  mutate(value = str_squish(value)) %>% 
  mutate(cluster = factor(cluster)) %>% 
  group_by(cluster, variable, value) %>% 
  tally() %>% 
  ungroup() %>% 
  group_by(cluster, variable) %>%  
  mutate(prop = n / sum(n) * 100)

categorical_plots_leva <- 
flakes_clustered_cats_long_leva %>% 
  ungroup(cluster) %>% 
  nest(-variable) %>% 
  mutate(plot = map(data , ~.x %>%  
                            ggplot(aes(cluster, 
                                          prop, 
                                          group=1, 
                                          fill = value)) + 
                            geom_col() +
                            xlab("") +
                           theme(legend.key.size = unit(0.4,"line"),
                                 legend.key.width = unit(0.2,"line"),
                                 legend.key.height = unit(0.4,"line"),
                                 legend.text = element_text(size=8))))

# get legend labels so we know what variable is plotted
# in each plot, not sure why I can't do this with map
for(i in 1:nrow(categorical_plots_leva)){
  categorical_plots_leva$plot[[i]] <- 
    categorical_plots_leva$plot[[i]] + guides(fill=guide_legend(title=unique(categorical_plots_leva$variable)[i]))
}

# arrange and draw the plots
cowplot::plot_grid(plotlist = categorical_plots_leva$plot,
                   align = "hv",
                   axis = "tblr")

```